RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2014-08-25 manuscripts/manuscript.Rmd
update get_characters / get_characters_list Adds option to use proper column instead of rownames for the taxa, as suggested in #89.  Closes #89

---
#layout: preprint
layout: review, 11pt
title: "RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: ropensci
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: ropensci
    address: |
      University of California, Berkeley, CA, USA
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands
abstract: | 
      1. NeXML is a new phylogenetic informatics data and metadata exchange 
      format that offers substantial improvements in reliability, extensibility, 
      <!-- refer to NeXML as a format and NEXUS as a standard, correct? -->
      and richness over earlier standards such as NEXUS. In this paper we present 
      the RNeXML package which provides a bridge between this rich and reliable
      data format and the extensive library of phylogenetic tools already
      available in R, with a particular emphasis on comparative methods.

@@

The `taxize_nexml()` function uses the R package `taxize` 
[@Chamberlain_2013] to check each taxon label against the NCBI database.
If a unique match is found, a metadata annotation is added to the taxon
providing the NCBI identification number to the taxonomic unit.  If no
match is found, the user is warned to check for possible typographic
errors in the taxonomic labels provided.

```{r message=FALSE, results='hide'}
birds <- taxize_nexml(birds, "NCBI")
```

<!-- what if multiple NCBI taxon IDs are found, e.g. 'Aotus'? -->

>>>>

---
#layout: preprint
layout: review, 11pt
title: "RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: ropensci
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: ropensci
    address: |
      University of California, Berkeley, CA, USA
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands
abstract: | 
      1. NeXML is a new phylogenetic informatics data and metadata exchange 
      format that offers substantial improvements in reliability, extensibility, 
      and richness over earlier formats such as NEXUS. In this paper we present 
      the RNeXML package which provides a bridge between this rich and reliable
      data format and the extensive library of phylogenetic tools already
      available in R, with a particular emphasis on comparative methods.

@@

The `taxize_nexml()` function uses the R package `taxize` 
[@Chamberlain_2013] to check each taxon label against the NCBI database.
If a unique match is found, a metadata annotation is added to the taxon
providing the NCBI identification number to the taxonomic unit.

```{r message=FALSE, results='hide'}
birds <- taxize_nexml(birds, "NCBI")
```
If no match is found, the user is warned to check for possible typographic
errors in the taxonomic labels provided. If multiple matches are found,
the user will be prompted to choose between them.
